Replication and Extension of Analyses Regarding Effects of Highly Restrictive Abortion Policies

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@

The main limitations of these analyses are due to missing data. Both the original analysis and our replication of this analysis are limited by the use of data from only 18 states.  We addressed this shortcoming in our extension, though our extension still does not have data for all states and/or all counties. Further, our replication of the original analysis as well as our extensions and expansions of said analysis are limited by the fact that state-level legislative data is only available for the years 2000, 2010, and 2020. More detailed legislative information is available for more recent years (roughly 2018 to the present), but abortion count data for these years is not yet released. Further analyses with more recent, detailed legislative data and abortion counts from more recent years would greatly add to the present analyses and could also allow for a more longitudinal approach. The original analysis and our own are also limited by the fact that abortion rates are calculated using total number of women, rather than number of women of reproductive age - this is due to the lack of data on population statistics stratified by age. Further, the data used in our extension is only by state, rather than by county, as previously discussed, due to suppression of data for the purpose of preserving anonymity.

>>>>

### Conclusions & Relevance
Our analysis confirms the results of Brown et al. showing a negative association between restrictive abortion legislation and abortion rates based on both the 18 states they analyized and based on data from all 50 states.  We furthered the analysis to additional outcomes and found positive associations between restrictions and infant death rates, while restrictions do not have a significant association with birth rates or maternal or fetal mortality rates after adjusting for distance to providers. Ultimately, this suggests that while highly restrictive legislation might significantly decrease abortion rates, as shown by Brown et al., such legislation does not actually significantly increase birth rates and actually increases infant death rates. This conclusion ultimately suggests that policies enforcing strict legislation on abortion restrictions would most likely be ineffective at increasing populations - while abortions decrease, births do not increase, suggesting that pregnancies may be terminated or prevented in other ways, and infant deaths actually increase. All together, highly restrictive legislation appears to have negative consequences on populations across the United States: such legislation reduces accessibility to important healthcare, may cause less-safe abortions that are not performed in medical clinics, and may result in more babies being born who are very sick.

@@

The main limitations of these analyses are due to missing data. Both the original analysis and our replication of this analysis are limited by the use of data from only 18 states.  We addressed this shortcoming in our extension, though our extension still does not have data for all states and/or all counties. Further, our replication of the original analysis as well as our extensions and expansions of said analysis are limited by the fact that state-level legislative data is only available for the years 2000, 2010, and 2020. More detailed legislative information is available for more recent years (roughly 2018 to the present), but abortion count data for these years is not yet released. Further analyses with more recent, detailed legislative data and abortion counts from more recent years would greatly add to the present analyses and could also allow for a more longitudinal approach. The original analysis and our own are also limited by the fact that abortion rates are calculated using total number of women, rather than number of women of reproductive age - this is due to the lack of data on population statistics stratified by age. Further, the data used in our extension is only by state, rather than by county, as previously discussed, due to suppression of data for the purpose of preserving anonymity. As a final note, the nature of abortion, mortality, fertility, and birth data is limited by the fact that not all of these occurrences are reported, and, if they are, not all ofthem are reported with complete accuracy (for example, it's possible that the abortion rates is much higher than those reported in the datasets and that women do not report all abortions) due to the sensitive nature of such data.
